\chapter{Axioms of Quantum Mechanics}
\label{aqm}

Formally a quantum system is defined by four axioms,
describing states,observables,dynamics, and probabilities.
\section{States}
    \label{states}

\begin{axiom}[States]
    \label{sts}
Physical state is a vector in complex separable Hilbert space $ \mathcal{H} $
\[ |\psi \rangle \in \mathcal{H} .\]
\end{axiom}
%description and discussion of states vectors rays etc

\section{Observables}
    \label{observables}

\begin{axiom}[Observables]
    \label{observs}
Observables are self-adjoint operators $ \hat A= \hat A^\dagger$ on $\mathcal H$. 
\end{axiom}
% discussion of observables SA HERM etc.

\section{Dynamics}
    \label{dynamics}
\begin{axiom}[Dynamics]
    \label{dyn}
    There is an observable $ \hat H$ (Hamiltonian),which generates time evolution
    (unitary group):
    \begin{equation}
       \label {evol}
 \hat U(t) = e^{-i \hat Ht/\hbar}.
    \end{equation}
\end{axiom}

    Evolution may be seen as action on observables (Heisenberg picture)
\begin{align}
   \hat A(t)&= \hat U(t) \hat A \hat U^{-1}(t), \label {heis1}\\
    \frac{d \hat A}{d t} &= \frac{i}{\hbar}[\hat H,\hat A].\label{heis2}
\end{align}
    Or as action on states (Schr\"odinger picture)
\begin{align}
    |\psi(t)\rangle&= \hat U^{-1}(t) |\psi \rangle, \label{schr1} \\
    \frac{d |\psi \rangle}{d t} &=-\frac{i}{\hbar} \hat H |\psi \rangle. \label{schr2}
\end{align}
%here about dynamics

\section{Probabilities}
\label{probabilities}
\begin{axiom}[Born rule]
    \label{born-rule}
    \[  B_\psi=\frac{\langle{\psi}|{\phi}\rangle}{\|{\psi}\| \|{\phi}\|}  \]
     is probability amplitude of transition between states. Corresponding 
     probability is $ P_\psi=|B_\psi|^2 $.
    Expectation value of an observable in a state 
    \begin{equation}
       \label {exp1}
    \langle \hat A \rangle_\psi = \frac{\langle \psi| \hat A |\psi \rangle} {\langle \psi|\psi \rangle}
    \end{equation}
    comprises all available information.
\end{axiom}
%here about prob,measurement, projection etc

Quantum mechanics admits multiple mathematically equivalent formulations together
with a wide spectrum of physical interpretations. While the formalism is empirically
unrivaled, its conceptual foundations remain unsettled. This review surveys 
the principal mathematical frameworks—wave mechanics, operator and algebraic 
formulations, path integrals, phase--space methods, and information--theoretic 
approaches—and analyzes how different interpretations assign physical meaning to
a shared mathematical core. Emphasis is placed on structural commonalities, 
the role of geometry and symmetry, modern measurement theory, and the distinction 
between representation, structure, and ontology. The aim is not to advocate a particular
interpretation, but to delineate which elements of quantum theory are fixed by mathematical 
consistency and empirical adequacy, and which remain matters of interpretational choice.

This review advances a structural realist perspective on quantum mechanics. 
On this view, the primary content of the theory lies not in its putative 
microscopic ontology but in the invariant mathematical structures common 
to all formulations. Different representations --- wave functions, operators,
path integrals, or quasi-probabilities --- are understood as coordinate descriptions
of an underlying abstract structure. Interpretations diverge precisely at the 
point where additional ontological commitments are introduced beyond 
this structure. From this standpoint, the persistence of interpretational 
plurality is not a deficiency of quantum theory but a reflection of its 
structural completeness.

The aim is not to advocate a particular interpretation, but to delineate which 
elements of quantum theory are fixed by mathematical consistency and empirical 
adequacy, and which reflect optional ontological commitments --- a distinction 
central to contemporary foundational analysis.

Quantum mechanics (QM) is characterized by an enduring tension between 
extraordinary predictive success and persistent conceptual ambiguity. 
From its inception, formally distinct approaches --- most notably wave 
mechanics and matrix mechanics--- were developed independently, 
only later to be unified within an abstract Hilbert space framework. 
Despite this mathematical unification, disagreement concerning the physical 
meaning of the formalism has never fully subsided.

A striking feature of quantum theory is the coexistence of two forms of plurality. 
On the one hand, there exist multiple mathematically equivalent formulations of the theory.
On the other hand, there exists a multiplicity of physical interpretations 
that agree on all empirical predictions while differing sharply in ontology,
epistemology, and explanatory aims.

This review adopts a structural perspective. Rather than tracing historical 
development in detail, we organize the subject around the invariant mathematical
core of quantum mechanics and examine how various formulations represent this structure.
We then survey the principal interpretational strategies and analyze how 
they map physical meaning onto the same formal framework. The guiding distinction 
throughout is that between representation, structure, and interpretation.
Clarifying this distinction is a central concern of the modern foundations of quantum mechanics.

\section{Core Mathematical Structure}

At an abstract level, non-relativistic quantum mechanics may be characterized by the following elements:

\begin{enumerate}
\item A complex Hilbert space $\mathcal{H}$.
\item Physical observables represented by self-adjoint operators,
    or more generally by positive operator--valued measures (POVMs).
\item Physical states represented by rays in $\mathcal{H}$ or 
    equivalently by density operators.
\item Time evolution described by a strongly continuous one-parameter 
    unitary group generated by a Hamiltonian operator.
\item Composite systems described by tensor products of Hilbert spaces.
\end{enumerate}

This abstract structure is common to all standard formulations. 
Differences between formulations reflect different choices of 
representation or emphasis rather than distinct physical content.

\section{Wave Mechanics (Schrödinger Formulation)}

In the Schrödinger formulation, states are represented by complex wave-functions
$\psi(x,t)$ defined on configuration space. Dynamics is governed by the 
Schrödinger equation
\begin{equation}
 i\hbar \partial_t \psi(x,t) = \hat{H} \psi(x,t),
\end{equation}
where $\hat{H}$ is a self--adjoint differential operator.

\subsection{Merits and Limitations}

Wave mechanics offers strong intuitive appeal, particularly for single-particle 
systems, and allows direct comparison with classical mechanics through semiclassical 
approximations. However, its configuration--space character becomes increasingly 
abstract for many--body systems. Moreover, gauge freedom and global phase invariance 
demonstrate that the wave function cannot be interpreted as a directly 
observable physical field.

\section{Operator and Hilbert Space Formulation}
The operator formulation, developed by Heisenberg, Born, Jordan, and von Neumann,
emphasizes observables as operators acting on an abstract Hilbert space.
\subsection{Canonical Commutation Relations}
The fundamental non-commutativity of observables,
\begin{equation}
 [\hat{x}, \hat{p}] = i\hbar,
\end{equation}
encodes the departure from classical mechanics and underlies uncertainty 
relations and quantum kinematics.

\subsection{Structural Advantages}
This formulation is representation-independent and naturally accommodates symmetries
through unitary group representations. It provides the conceptual foundation for 
spectral theory, quantum measurement theory, and quantum information.

\section{Algebraic and $C^*$-Algebraic Formulations}
In the algebraic approach, observables form a non-commutative $C^*$-algebra,
and states are defined as positive, normalized linear functionals on this algebra.

\subsection{Structural Emphasis}

This formulation is particularly well suited for systems with infinitely 
many degrees of freedom, clarifying the role of super-selection sectors, 
inequivalent representations, and locality. It plays a central role in quantum
statistical mechanics and quantum field theory.

\section{Path Integral Formulation}

Feynman’s path integral formulation expresses transition amplitudes as sums over histories,
\begin{equation}
 \langle x_f,t_f | x_i,t_i \rangle
 = \int \mathcal{D}[x(t)]\, e^{\frac{i}{\hbar} S[x(t)]},
\end{equation}
where $S$ is the classical action functional.

\subsection{Conceptual and Mathematical Status}

While mathematically subtle and not fully rigorous in general, the path integral 
is formally equivalent to the operator formulation and is indispensable in 
perturbation theory, semiclassical analysis, and quantum field theory. 
Conceptually, it emphasizes continuity between classical and quantum dynamics 
through the action principle.

\section{Phase Space and Quasi-Probability Formulations}

Phase space approaches, including the Wigner--Weyl--Moyal formalism, 
represent quantum states by quasi--probability distributions on classical phase space.

\subsection{Nonclassical Features}

The appearance of negative values in quasi-probability distributions highlights 
the irreducible departure of quantum theory from classical statistical mechanics,
while preserving a close structural analogy.

\section{Information-Theoretic and Operational Approaches}

Modern operational approaches treat quantum theory primarily as a framework
for predicting outcomes of measurements and transformations, often 
motivated by quantum information theory.

\subsection{Quantum States as Information}

In these approaches, quantum states are interpreted as encoding information,
knowledge, or rational betting commitments rather than 
objective microscopic properties.

\section{Measurement Theory}

Measurement occupies a distinctive role in quantum mechanics,
linking the formalism to empirical observation.

\subsection{Projective Measurements and POVMs}

The traditional projection postulate is generalized by POVMs, providing a 
flexible and operationally complete description of realistic measurements.

\subsection{Decoherence}

Environment-induced decoherence explains the suppression of interference between 
macroscopically distinct states and the emergence of effective classical probabilities. 
However, decoherence alone does not account for the selection of individual outcomes.

\section{Physical Interpretations}

\subsection{Copenhagen--Type Interpretations}

Copenhagen--type interpretations emphasize complementarity and the necessity 
of classical language for describing experimental outcomes. The quantum 
state is treated primarily as a predictive tool.

\section{Geometry and Symmetry}

Geometric formulations describe quantum mechanics as Hamiltonian flow on 
complex projective Hilbert space equipped with symplectic and metric structures.
Berry phases exemplify the physical significance of quantum geometry.

\section{Discussion: Structure Versus Interpretation}

Certain features --- Hilbert space structure, unitary dynamics, and probabilistic 
predictions --- are invariant across all formulations. By contrast, ontological 
commitments, the meaning of probability, and the status of the quantum state are 
interpretation--dependent. Recognizing this distinction clarifies the scope and 
limits of foundational debates.
\section{Structural Realism and Quantum Geometry}

Structural realism holds that the enduring content of successful physical theories resides
in their mathematical structure rather than in their specific ontological interpretations.
Quantum mechanics provides a particularly compelling case for this position.
Despite radical disagreement concerning the nature of quantum states, 
measurement, and reality, all viable interpretations preserve the same core formal structure.

From a geometric perspective, quantum mechanics may be formulated as a Hamiltonian theory
on complex projective Hilbert space. Physical pure states correspond to points in projective
space, observables generate Hamiltonian flows, and transition probabilities are encoded 
in the Fubini--Study metric. Within this framework, global phase invariance and gauge 
freedom are not interpretational artifacts but structural necessities.

Geometric phases, most notably the Berry phase, further reinforce the structural viewpoint.
Such phases depend only on the geometry of parameter space and not on the detailed dynamics
or representation of the system. Their observable consequences demonstrate that relational 
and topological features of the formalism possess direct physical significance, independent
of any underlying microscopic ontology.

From the structural realist standpoint, attempts to reify particular representational elements
---such as wave functions in configuration space or particle trajectories ---
go beyond what the formalism itself warrants. What quantum mechanics robustly represents
is a network of relations encoded in its algebraic and geometric structures. Ontological
narratives that extend beyond this structure are interpretational overlays rather than 
consequences of the theory.
\section{Discussion: Structure Versus Ontology}

The comparison of formulations and interpretations suggests a clear division between 
structural and interpretational content. Hilbert space structure, non-commutativity of 
observables, unitary dynamics, and probabilistic predictions are invariant across all
formulations of quantum mechanics. These elements form the stable core of the theory 
and are preserved under all empirically viable interpretations.

By contrast, questions concerning the reality of the wave function, the existence of
multiple worlds, or the determinism of particle trajectories depend on additional 
metaphysical assumptions. Such assumptions are not fixed by the formalism itself. 
Structural realism therefore provides a principled explanation for the persistence 
of interpretational plurality: different interpretations correspond to distinct 
ontological extensions of the same underlying structure.

From this perspective, foundational debates are best understood not as disputes
about empirical content, but as disagreements about how—or whether—to supplement 
the mathematical structure of quantum mechanics with further ontological claims.

\section{Conclusion}

Quantum mechanics is best understood as a stable mathematical structure 
admitting multiple representations and interpretations. Progress in foundations
is likely to arise not from selecting a uniquely correct interpretation, 
but from clarifying which questions are meaningful within the theory’s structural
constraints and which lie beyond them.
Viewed through a structural realist lens, quantum mechanics emerges as a 
theory whose physical content is exhausted by its mathematical and geometric
structure. The multiplicity of formulations reflects representational freedom, 
while the diversity of interpretations reflects optional ontological commitments. 
Recognizing this distinction clarifies the scope of foundational inquiry and 
suggests that the enduring lesson of quantum mechanics is not the failure of 
realism per se, but the primacy of structure over substance in fundamental physics.
From a structural realist perspective, the wave function is best understood not as 
a physical field inhabiting configuration space, but as a representational device 
encoding the geometry and algebraic relations that constitute the true invariant 
content of quantum theory.

\appendix
\section{Geometric Structure of Projective Hilbert Space}

The physical state space of pure quantum states is not the Hilbert space 
$\mathcal{H}$ itself, but the associated complex projective space 
$\mathbb{P}(\mathcal{H})$, obtained by identifying vectors that differ 
by a nonzero complex scalar. This identification reflects the physical 
irrelevance of global phase and normalization.

Projective Hilbert space carries a natural Kähler structure, combining a 
symplectic form, a Riemannian metric, and a compatible complex structure.
The symplectic form underlies the Hamiltonian character of quantum dynamics, 
while the Riemannian structure is given by the Fubini--Study metric. 
Transition probabilities between pure states are directly related to 
geodesic distances in this metric.

Within this geometric framework, observables correspond to Hamiltonian 
vector fields generating unitary flows on $\mathbb{P}(\mathcal{H})$. 
Schrödinger evolution is thereby reinterpreted as Hamiltonian motion 
on a curved state space, placing quantum mechanics firmly within the 
general framework of geometric mechanics.

Geometric phases provide a striking illustration of the physical 
significance of this structure. The Berry phase depends solely on the 
holonomy induced by parallel transport in projective Hilbert space and 
is independent of dynamical details. Its observability demonstrates that 
relational and topological features of the quantum state space possess
direct physical meaning.

From a foundational perspective, the projective geometry of quantum 
state space reinforces a structural realist interpretation. Physical 
content is encoded in invariant geometric relations rather than in 
representational artifacts such as wave functions in configuration space. 
The geometry of $\mathbb{P}(\mathcal{H})$ thus provides a natural arena in
which the mathematical structure of quantum mechanics acquires clear physical
significance without requiring additional ontological postulates.
\subsection*{Gauge Structure and Representational Redundancy}

Hilbert space provides a linear representation of quantum states, 
but this linearity exceeds physical necessity. The physically relevant state
space is obtained by quotienting $\mathcal{H} \setminus \{0\}$ by the action 
of the complex multiplicative group. This quotient structure is not optional:
it is enforced by the empirical indistinguishability of states differing by 
global phase or normalization.

From this standpoint, wave function realism resembles a gauge-fixed ontology. 
Just as electromagnetic potentials acquire physical meaning only up to 
gauge equivalence, wave functions acquire physical significance only through 
equivalence classes. Ontological commitment to the wave function itself therefore 
parallels commitment to a particular gauge choice, 
rather than to a gauge-invariant physical entity.

This analogy is sharpened by geometric formulations of quantum mechanics, 
in which the fundamental objects are points and curves in projective Hilbert
space. In this setting, the wave function appears only as a local section of 
a fiber bundle, while all observable quantities are functions of gauge-invariant
geometric structures. Treating the wave function as ontologically fundamental 
thus conflicts with the natural geometric organization of the theory.

\chapter{Quantizations }
\label{qnt}

\section{Deformation Quantization}

Deformation quantization replaces the pointwise product of functions on phase space by a star product
\begin{equation}
f \star g = fg + \frac{i\hbar}{2}{f,g} + O(\hbar^2).
\end{equation}
Quantum mechanics is then viewed as a deformation of the classical observable algebra rather than as an operator theory ab initio.

\section{Geometric Quantization}

Geometric quantization begins with a complex line bundle $L \to M$ equipped with a connection $\nabla$ satisfying
\begin{equation}
\mathrm{curv}(\nabla) = \frac{1}{\hbar}\omega.
\end{equation}
This requires the integrality condition $[\omega]/2\pi\hbar \in H^2(M,\mathbb{Z})$.

To obtain a physical Hilbert space, one introduces a polarization selecting a subset of compatible observables. The choice of polarization is generally non-unique and may not exist globally.

\section{Geometric Quantization}

Deformation quantization modifies the algebra. 
Geometric quantization instead attempts to construct the Hilbert space 
directly from symplectic geometry.

\subsection{Prequantization}

Given $(M,\omega)$ with integral symplectic form:

\[
[\omega] \in H^2(M,\mathbb Z),
\]

there exists a Hermitian line bundle $L \to M$ with curvature:

\[
F_\nabla = -i\omega.
\]

Prequantum Hilbert space:

\[
\mathcal H_{pre} = L^2(M,L).
\]

Prequantum operator:

\[
\hat f = -i\hbar \nabla_{X_f} + f.
\]

However:

\[
[\hat f, \hat g] = i\hbar \widehat{\{f,g\}}.
\]

This space is too large.

\subsection{Polarization}

To reduce degrees of freedom, choose a polarization 
(e.g., vertical polarization in $T^*Q$). 

Wavefunctions become sections constant along half of phase space directions.

Example:
\[
\mathcal H = L^2(Q).
\]

\subsection{Metaplectic Correction}

For consistency of half-forms, one introduces the metaplectic correction.

\subsection{Non-Functoriality}

Quantization is not a functor:

\[
\text{Symplectic manifolds}
\not\longrightarrow
\text{Hilbert spaces}
\]

in a structure-preserving categorical sense.

Obstructions include:

\begin{itemize}
\item Groenewold–Van Hove no-go theorem
\item Dependence on polarization
\item Global topological constraints
\end{itemize}

Thus quantization is not canonical.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Path Integral Quantization}

In the path-integral approach, transition amplitudes are formally expressed as
\begin{equation}
\langle q_f,t_f | q_i,t_i \rangle = \int \mathcal{D}q, e^{\frac{i}{\hbar}S[q]},
\end{equation}
where $S$ is the classical action. While powerful and intuitive, this formulation typically requires auxiliary constructions for mathematical rigor.

\section{Stochastic quantization}
\label{stochastic-quantization}

\section{Tomographic Quantization}
\label{tomographic-quantization}

