\chapter{Canonical Quantization Examples}
\label{cqe}

For most practical problems canonical quantization is realized as
\begin{equation}
q_i \mapsto \hat q_i = x_i, \qquad p_j \mapsto \hat p_j = -i\hbar \frac{\partial}{\partial x_j},
\end{equation}
acting on $\mathcal{H}=L^2(\mathbb{R}^n,dx)$. \footnote{ Actually this representation
is essentially unique.}
State in this representation will be called "wave function".

Below we shall consider a number of very simple but instructive 
examples, showing how non-locality of wave function contradicts classical
physical intuition, trained on classical point particles.

\section{The Ground State Determines  Dynamics }
\label{cqe:gsdd}

In classical mechanics, dynamics is specified by a Hamiltonian function;
individual solutions do not determine the Hamiltonian. In quantum
mechanics, however, the structure of the Schr\"odinger equation implies
a striking inversion: knowledge of a single, sufficiently regular
stationary state---specifically the nondegenerate ground state---fixes
the potential and thereby all dynamics.

Consider a single nonrelativistic particle of mass $m$ on $\mathbb{R}^n$ governed by
\begin{equation}
\hat H = -\frac{\hbar^2}{2m}\nabla^2 + V(\vec x),
\end{equation}
with $V$ real-valued, locally bounded, and confining enough to admit a 
normalizable ground state $\psi_0>0$ almost everywhere.

We assume:
\begin{enumerate}
\item The ground state is nondegenerate.
\item $\psi_0(\vec{x})$ is twice differentiable and nowhere vanishing.
\item Boundary conditions (e.g. decay at infinity) are known.
\end{enumerate}
These assumptions are standard in both mathematical physics and quantum chemistry.

Let $E_0$ be the ground-state energy. The time-independent Schr\"odinger equation reads
\begin{equation}
-\frac{\hbar^2}{2m}\nabla^2\psi_0 + V\psi_0 = E_0\psi_0.
\end{equation}
Solving for the potential yields
\begin{equation}
    \label{pot1d}
V(\vec{x}) = E_0 + \frac{\hbar^2}{2m}\frac{\nabla^2\psi_0(\vec{x})}{\psi_0(\vec{x})}.
\end{equation}
Thus, given $\psi_0$ and $E_0$, the potential is uniquely determined
pointwise. Since adding a constant to $V$ merely shifts all energies
uniformly, the potential is fixed up to an additive constant even if
$E_0$ is unknown.

The positivity of $\psi_0$ ensures that the reconstruction formula is
well defined. Any other potential $V'$ admitting the same ground 
state must coincide with $V$ almost everywhere. Hence the ground 
state uniquely specifies the Hamiltonian.

Once $V$ is fixed, the entire spectral problem is fixed: all excited
eigenstates and eigenvalues are determined, the propagator 
$U(t)=e^{-i\hat H t/\hbar}$ is fixed, and all expectation values
and transition amplitudes follow. Therefore, the ground state 
determines not merely low-energy physics but the complete time 
evolution of the system.

In one dimension, the result is a special case of inverse 
Sturm--Liouville theory, where knowledge of the lowest 
eigenfunction suffices to reconstruct the potential locally.

Defining the superpotential
\begin{equation}
W(x) = -\frac{\hbar}{\sqrt{2m}}\frac{\psi_0'(x)}{\psi_0(x)},
\end{equation}
we obtain
\begin{equation}
V(x) = W^2(x) - \frac{\hbar}{\sqrt{2m}}W'(x) + E_0,
\end{equation}
showing again that $\psi_0$ encodes the full Hamiltonian structure.


Let the ground-state wave function be written as
\begin{equation}
\psi_0(\vec{x}) = e^{-\phi(\vec{x})},
\end{equation}
where $\phi$ is a smooth real-valued function. Positivity of the ground state 
guarantees that this representation is globally well defined. Substituting 
into the reconstruction formula yields
\begin{equation}
V(\vec{x}) = E_0 + \frac{\hbar^2}{2m}\Big[(\nabla \phi)^2 - \nabla^2 \phi\Big].
\end{equation}
The potential is thus a scalar functional constructed from first 
and second derivatives of $\phi$, suggesting an intrinsic geometric interpretation.

Degenerate ground states may lead to non-uniqueness. Excited states 
cannot be used directly due to the presence of nodes. For interacting
many-body systems the statement generalizes formally, but 
reconstruction becomes non-local in configuration space. 

This determinacy challenges the intuition that dynamics is prior to 
states. In the Schr\"odinger framework, the ground state contains 
the full dynamical content. The Hamiltonian may therefore be regarded
as a derived object, secondary to a privileged stationary state. 
This perspective aligns naturally with structural and geometric 
interpretations of quantum theory.

\subsection{Explicit Example}

Consider the nontrivial ground state
\begin{equation}
    \label{gsx4}
\psi_0(x) = A \exp\Big(-\frac{a}{4} x^4\Big), \quad a>0.
\end{equation}

Substituting \eqref{gsx4} into \eqref{pot1d}, we compute:
\begin{align*}
\psi_0'(x) &= -a x^3 \psi_0(x), \
\psi_0''(x) &= (a^2 x^6 - 3 a x^2) \psi_0(x).
\end{align*}
Thus the potential is
\begin{equation}
V(x) = E_0 + \frac{\hbar^2}{2m} (a^2 x^6 - 3 a x^2),
\end{equation}
a sextic potential with a negative quadratic term, uniquely determined by $\psi_0$.

Writing excited states as $\psi_n = \psi_0 P_n$, the Sturm--Liouville operator for $P_n$ is
\begin{equation}
L = -\frac{\hbar^2}{2m} \left( \frac{d^2}{dx^2} + 2 a x^3 \frac{d}{dx} \right),
\end{equation}
with weight $w(x) = \psi_0^2 = A^2 e^{- \frac{a}{2} x^4}$. The eigenvalue equation is
\begin{equation}
L P_n = (E_n - E_0) P_n,
\end{equation}
which defines the spectrum and orthogonal polynomials (Freud polynomials) relative to $w(x)$.
Numerical or recursive methods can be used to compute $E_n$ and $P_n$.

\section{Hard Wall}
\label{cqe:hw}
Let us consider a particle moving at half line:
\begin{equation}
    \label{cqe:hwp}
V(x) = 
\begin{cases}
0, & 0 < x ,\\
\infty, & x<0.
\end{cases}
\end{equation}

The time–independent Schrödinger equation is
\[
-\frac{\hbar^2}{2m}\psi''(x)
= E\psi(x).
\]
Define $ k^2=\frac{2mE}{\hbar^2} $,then
\begin{equation}
    \label{cqe:hwe1}
\psi''(x)+k^2 \psi(x)=0.
\end{equation}
Surely we should add boundary condition $ \psi(0)=0 $.
The solution (which can not be normalized) is evidently 
\begin{equation}
   \label {cqe:hws1}
    \psi(x)=\sin(k x)
\end{equation}
Physically the situation is clear: interference of in-going and out-going waves, 
however since values of $x$ are restricted $ x \ge 0 $,
no self-adjoint operator of momentum satisfying canonical commutation
relations \eqref{cq:ccr} exists \footnote{ Just here mathematically oriented 
people start talking about difference between Hermitian and 
self-adjoint operators}. So momentum is not observable--- rather 
strange statement!

Physical answer is based on remark that potential \eqref{cqe:hwp} 
is only idealization used to simplify calculations.

We should start from some smooth but fast growing potential,
like
\begin{equation}
   \label {cqe:swp}
   V(x)=g/x^2,\quad g>0.
\end{equation}
So time–independent Schrödinger equation becomes
\begin{equation}
    \label{cqe:swe1}
\psi''(x)+\left(k^2-\frac{\gamma}{x^2}\right)\psi(x)=0.
\end{equation}
where 
\[
k^2=\frac{2mE}{\hbar^2},
\quad
\gamma=\frac{2mg}{\hbar^2}.
\]
Let $\nu=\sqrt{\frac{1}{4}+\gamma}$.
The general solution for $E>0$ finite at small $x$ is 
\begin{equation}
    \label{cqe:sws}
\psi(x)=\sqrt{x}A J_\nu(kx).
\end{equation}
\begin{remark}
For $g>0$, the Hamiltonian is essentially self–adjoint.
No boundary condition at $x=0$ is imposed externally.
\end{remark}
As $x\to 0$,
\[\psi(x)\sim x^{\nu+\frac{1}{2}}\]
and ass $x \to \infty$
\[
J_\nu(kx)
\sim
\sqrt{\frac{2}{\pi kx}}
\cos\left(
kx-\frac{\pi\nu}{2}-\frac{\pi}{4}
\right).
\]
When $g \to 0$ $\nu \to 1/2$ since
\[J_{1/2}(x)=sqrt{\frac{2}{\pi x}} \sin(x)\]
the solution of \eqref{cqe:swe1} becomes that of \eqref{cqe:hwe1},
without need in boundary conditions.

\section{Particle in Box}

Consider a particle confined to a one-dimensional infinite square well of width $L$,
\begin{equation}
V(x) = 
\begin{cases}
0, & 0 < x < L,\\
\infty, & \text{otherwise}.
\end{cases}
\end{equation}

A \emph{wave packet} is a superposition of stationary states,
\begin{equation}
\Psi(x,0) = \sum_{n=1}^{\infty} c_n \, \phi_n(x), 
\qquad
\phi_n(x) = \sqrt{\frac{2}{L}} \sin \frac{n \pi x}{L},
\end{equation}
where the coefficients $c_n$ encode the initial localization and momentum distribution of the particle.  
Time evolution follows from the Schr\"odinger equation,
\begin{equation}
\Psi(x,t) = \sum_{n=1}^{\infty} c_n \, \phi_n(x) \, e^{-i E_n t/\hbar}, 
\qquad
E_n = \frac{n^2 \pi^2 \hbar^2}{2 m L^2}.
\end{equation}
The quadratic energy spectrum is responsible for the characteristic quantum features described below.

For a wave packet narrowly localized in both position and momentum (centered around a large $n_0$):
\begin{itemize}
    \item The packet moves approximately according to the classical trajectory,
    \begin{equation}
    x_{\mathrm{cl}}(t) = x_0 + v t,
    \end{equation}
    reflecting elastically at the walls.
    \item Ehrenfest's theorem ensures that the expectation values of position and momentum follow classical equations of motion at short times.
\end{itemize}

Due to the quadratic spectrum $E_n \sim n^2$, different stationary components acquire phases at different rates.  
Consequences include:
\begin{itemize}
    \item Spreading (dispersion) of the wave packet.
    \item Destruction of the initial localization by interference.
    \item A characteristic \emph{collapse time} after which the probability density appears irregular.
\end{itemize}
The evolution remains fully unitary; no energy loss or dissipation occurs.

The exact quadratic spectrum leads to wave packet \emph{revivals}:
\begin{equation}
T_{\mathrm{rev}} = \frac{4 m L^2}{\pi \hbar}.
\end{equation}
At $t = T_{\mathrm{rev}}$, the wave packet reconstructs itself up to a global phase,
\begin{equation}
\Psi(x,T_{\mathrm{rev}}) = e^{i \theta} \Psi(x,0),
\end{equation}
while at rational fractions $t = \frac{m}{k} T_{\mathrm{rev}}$, \emph{fractional revivals} occur, splitting the packet into $k$ localized copies.


The dynamics of a wave packet in a square well illustrate:
\begin{description}
    \item [Short-time classical motion:] The packet moves and reflects approximately like a classical particle.
    \item [Quantum dispersion:] Interference between stationary components spreads the packet.
    \item [Exact revivals:] The wave packet periodically reconstructs itself due to the quadratic spectrum.
    \item [Unitary boundary-induced interference:] Reflections at the walls generate characteristic oscillations without dissipation.
\end{description}
This system exemplifies the interplay between classical intuition and fundamentally quantum effects.
However despite the fact that packet behaves as if momentum reverses sign at the walls,
no self-adjoint momentum operator exists. So momentum is not observable!

Let us again go to smooth potential.
\[
    V(x)=a^2\cot^2(bx), \qquad x\in(0,\pi/b),
\]
The Schr\"odinger equation is 
\begin{equation}
    \label{cqe:pte}
    \frac{\hbar^2}{2m}\frac{d^2 \Psi}{dx^2}+(E-a^2\cot^2(bx))\Psi
\end{equation}
Let 
\[ k^2=\frac{2 m (E+a^2)}{\hbar^2 b^2} \quad 
\lambda(\lambda-1)=\frac{2 m a^2}{\hbar^2 b^2}
\]
so equation \eqref{cqe:pte} becomes
\[ \Psi''+(k^2-\frac{\lambda (\lambda-1)}{\sin^2(x)}) \Psi=0 \]
Substituting \[ \Psi(x)=\sin^\lambda(x) y(\cos(x)) \] we get
equation ($ \xi=\cos(x) $)
\[ y''-\frac{(2 \lambda+1)\xi}{1-\xi^2} y' +\frac{n(n+2 \lambda)}{1-\xi^2}y=0 , \] where $ k^2-\lambda^2=n(n+2 \lambda) $
Gegenbauer polynomials $G^\lambda_n$ are solutions of this equation.
So eigenvalues are 
\begin{equation}
   \label {cqe:ptev}
   E_n=\frac{\hbar ^2 b^2}{2 m}(n^2+(2 n+1) \lambda)
\end{equation}
and eigenfunctions are 
\begin{equation}
   \label {cqe:ptef}
    \Psi_n=A_n \sin^\lambda(x) G^\lambda_n(cos(x)) \quad x\in (0,\pi)
\end{equation}
Let \[ \gamma=\frac{8 m a^2}{\hbar^2 b^2} \] then \[ 
\lambda=\frac{1}{2} (1+\sqrt{1+\gamma})\]
Since $ \lambda>1 $ and $ G^\lambda_n $ are finite, one can see that there is no need in boundary conditions.

Note that $\lambda \to 1$ when $ \gamma \to 0 $ and \[ 
G^1_n(\cos x)=\frac{\sin(n+1)x}{\sin(x)}\] so eigenvalues \eqref{cqe:ptev}
and eigenfunctions \eqref{cqe:ptef} become that of the box.

\begin{remark}
    Actually we obtain Schr\"odinger equation without explicit reference to momentum 
    operator. Kinetic energy becomes Laplacian operator, potential acts as multiplication.
    Explicit use of momentum becomes necessary when magnetic field is included.
\end{remark}

\section{Constraint Quantization}
\label{cqe:cq}

Quantum mechanics on curved manifolds arises naturally when a particle
is constrained to move on a surface embedded in $\mathbb{R}^3$.
A physically consistent derivation of the effective surface Hamiltonian 
was given by da Costa using a thin-layer confining procedure. The key 
result is the appearance of an additional curvature-induced potential 
term depending on the mean and Gaussian curvatures.

\subsection{Constraint Classification: First-Class vs Second-Class}
In Hamiltonian mechanics, constraints $\phi_i(q,p)=0$ are classified 
according to their Poisson bracket algebra.

A constraint $\phi$ is called first-class if its Poisson brackets with 
all constraints vanish on the constraint surface:
\begin{equation}
{\phi_i,\phi_j} \approx 0.
\end{equation}
(Here $\approx$ denotes equality on the constraint surface.)
First-class constraints generate gauge transformations. The reduced phase 
space is obtained by quotienting the constraint surface by the gauge orbits.
This is the setting of Marsden--Weinstein reduction and underlies many 
gauge theories and systems with symmetries.

Constraints are second-class if the matrix
\begin{equation}
C_{ij} = {\phi_i,\phi_j}
\end{equation}
is invertible on the constraint surface. In this case, the constraints do 
not generate gauge symmetries. Instead, one defines the Dirac bracket
\begin{equation}
{A,B}_D = {A,B} - {A,\phi_i}C^{ij}{\phi_j,B},
\end{equation}
which ensures that all constraints hold strongly (not merely weakly).

The surface-constrained particle belongs to this category: the position 
constraint $F(\mathbf{x})=0$ and its consistency condition 
$\nabla F \cdot \mathbf{p}=0$ form a second-class pair. No gauge symmetry 
is involved, and reduction proceeds via Dirac brackets rather than quotient
by group action.
\subsection{Thin-Layer Quantization}

The effective Hamiltonian takes the form
\begin{equation}
H_{\text{eff}} = -\frac{\hbar^2}{2m}\Delta_{LB} - \frac{\hbar^2}{2m}(H^2 - K),
\end{equation}
where $\Delta_{LB}$ is the Laplace--Beltrami operator, $H$ the mean curvature,
and $K$ the Gaussian curvature of the surface.

Let $\Sigma$ be a smooth surface embedded in $\mathbb{R}^3$. Introduce adapted 
coordinates $(q^1,q^2,u)$ where $(q^1,q^2)$ parametrize the surface and $u$ is 
the normal coordinate. The embedding is
\begin{equation}
\mathbf{r}(q^1,q^2,u) = \mathbf{R}(q^1,q^2) + u \mathbf{n}(q^1,q^2),
\end{equation}
with $\mathbf{n}$ the unit normal.

The metric near the surface becomes
\begin{equation}
G_{ab}(u) = g_{ab} - 2u b_{ab} + u^2 (b g^{-1} b)*{ab},
\end{equation}
where $g*{ab}$ is the surface metric and $b_{ab}$ the second fundamental form.

After introducing a strong confining potential $V(u)$ and separating the 
normal mode, the effective Hamiltonian on the surface becomes
\begin{equation}
H_{\text{eff}} = -\frac{\hbar^2}{2m}\Delta_{LB} - \frac{\hbar^2}{2m}(H^2 - K).
\end{equation}

For first-class systems, reduction corresponds to selecting gauge-invariant 
states, and geometric quantization techniques often apply. For second-class
systems, reduction modifies the fundamental bracket structure prior to 
quantization. The thin-layer procedure effectively implements a quantum-level
analogue of eliminating second-class constraints, leading to curvature-induced 
terms absent in purely classical Dirac reduction.

This distinction clarifies why the geometric potential arises in constrained 
surface quantum mechanics but does not contradict the Guillemin--Sternberg theorem.

\section{Quantum Particle Constrained to a Torus}

Consider a standard torus embedded in $\mathbb{R}^3$:
\begin{equation}
\begin{aligned}
x &= (R + r \cos\theta) \cos\phi, \
y &= (R + r \cos\theta) \sin\phi, \
z &= r \sin\theta,
\end{aligned}
\end{equation}
where:
\begin{itemize}
\item $R$ = major radius
\item $r$ = minor radius
\item $0 \le \theta, \phi < 2\pi$
\end{itemize}

The induced metric is diagonal:
\begin{equation}
ds^2 = r^2 d\theta^2 + (R + r \cos\theta)^2 d\phi^2.
\end{equation}
Thus
\begin{equation}
g_{\theta\theta} = r^2, \qquad g_{\phi\phi} = (R + r\cos\theta)^2.
\end{equation}

The principal curvatures are
\begin{equation}
k_1 = \frac{\cos\theta}{r},
\qquad
k_2 = \frac{1}{R + r \cos\theta}.
\end{equation}
Hence
\begin{equation}
H = \frac{1}{2}(k_1 + k_2),
\qquad
K = k_1 k_2.
\end{equation}
Explicitly,
\begin{equation}
H = \frac{1}{2}\left(\frac{\cos\theta}{r} + \frac{1}{R + r \cos\theta}\right),
\end{equation}
\begin{equation}
K = \frac{\cos\theta}{r(R + r \cos\theta)}.
\end{equation}

The geometric potential is
\begin{equation}
V_g = -\frac{\hbar^2}{2m}(H^2 - K).
\end{equation}
Compute
\begin{equation}
H^2 - K = \frac{1}{4}(k_1 - k_2)^2.
\end{equation}
Thus
\begin{equation}
V_g = -\frac{\hbar^2}{8m}(k_1 - k_2)^2.
\end{equation}
Substituting the torus curvatures:
\begin{equation}
V_g(\theta) = -\frac{\hbar^2}{8m}
\left(
\frac{\cos\theta}{r} - \frac{1}{R + r \cos\theta}
\right)^2.
\end{equation}
This is the explicit curvature-induced potential.

Unlike the sphere, for which $H^2 - K = 0$, the torus has nontrivial 
curvature variation. The geometric potential is attractive and depends 
only on $\theta$. It is strongest on the inner equator ($\theta=\pi$) 
and weakest on the outer equator ($\theta=0$).
Thus curvature alone generates an effective quantum potential even 
in absence of external forces.

\subsection{Dirac-Bracket vs Thin-Layer Quantization}

Two conceptually distinct but related procedures lead to quantum 
mechanics on a curved surface:
Dirac-Bracket Quantization begins with the classical constrained Hamiltonian 
system in $T^*\mathbb{R}^3$ defined by the second-class constraints
\begin{equation}
F(\mathbf{x}) = 0, \qquad \nabla F \cdot \mathbf{p} = 0.
\end{equation}
The Poisson brackets are replaced by Dirac brackets
\begin{equation}
{A,B}*D = {A,B} - {A,\phi_i}C^{ij}{\phi_j,B},
\end{equation}
where $C^{ij}$ is the inverse constraint matrix. This projects momenta onto 
the tangent bundle $T\Sigma$ and produces a reduced classical Hamiltonian
\begin{equation}
H*{cl} = \frac{1}{2m} g^{ab} p_a p_b.
\end{equation}
Quantization is then performed by promoting Dirac brackets to commutators. 
One obtains formally
\begin{equation}
\hat H = -\frac{\hbar^2}{2m}\Delta_{LB},
\end{equation}
with no additional potential term appearing at the purely algebraic level.

In the thin-layer method one begins directly with the three-dimensional 
Schr"odinger equation and introduces a strong confining potential in the
normal direction. After separating the transverse ground state and taking
the limit of vanishing layer thickness, one obtains
\begin{equation}
H_{eff} = -\frac{\hbar^2}{2m}\Delta_{LB} - \frac{\hbar^2}{2m}(H^2-K).
\end{equation}
The extra curvature term arises from the Jacobian factor $\sqrt{G}$ of the
three-dimensional metric and the elimination of the normal coordinate in 
the kinetic operator.

Classically, both approaches produce the same reduced phase space $T^*\Sigma$ 
and identical Hamiltonian flow. The difference emerges at the quantum level:
\begin{itemize}
\item Dirac reduction quantizes the reduced classical algebra directly.
\item Thin-layer quantization reduces the quantum operator itself.
\end{itemize}

Because the Laplacian depends on the full embedding metric, eliminating the 
transverse direction before or after quantization is not equivalent. The term
$(H^2-K)$ therefore represents a quantum correction generated by 
noncommutativity between reduction and quantization:
\begin{equation}
\text{Quantize}\circ\text{Reduce} \neq \text{Reduce}\circ\text{Quantize}.
\end{equation}

For the sphere the correction vanishes identically, while for the torus it
produces the nontrivial geometric potential derived above.

\subsection{Noncommutativity of Quantization and Reduction}

We now formalize the structural statement underlying the appearance of 
the geometric potential.
\begin{theorem}[Noncommutativity of Quantization and Reduction]
Let $M = T^*\mathbb{R}^3$ with canonical symplectic form and let 
    $\Sigma \subset \mathbb{R}^3$ be a smooth embedded surface defined
    by a regular constraint $F(\mathbf{x})=0$. Denote by $R$ the classical
    symplectic reduction map
\begin{equation}
R: T^*\mathbb{R}^3 \longrightarrow T^*\Sigma,
\end{equation}
and by $Q$ a canonical quantization procedure mapping classical 
    Hamiltonians to self-adjoint operators.

Then, for the free Hamiltonian $H = \frac{\mathbf{p}^2}{2m}$,
\begin{equation}
Q(R(H)) \neq R(Q(H))
\end{equation}
in general. More precisely,
\begin{align}
Q(R(H)) &= -\frac{\hbar^2}{2m}\Delta_{LB}, 
R(Q(H)) &= -\frac{\hbar^2}{2m}\Delta_{LB} - \frac{\hbar^2}{2m}(H^2-K),
\end{align}
where $H$ and $K$ denote the mean and Gaussian curvatures of $\Sigma$.
\end{theorem}

\begin{proof}[Sketch of Proof]
The classical reduction projects momenta onto the tangent bundle and 
    yields the reduced Hamiltonian $H_{cl} = \frac{1}{2m} g^{ab} p_a p_b$.
    Canonical quantization of this expression produces 
    the Laplace--Beltrami operator.

Conversely, quantizing first yields the three-dimensional Laplacian
\begin{equation}
\hat H = -\frac{\hbar^2}{2m} \Delta_{\mathbb{R}^3}.
\end{equation}
Expressing $\Delta_{\mathbb{R}^3}$ in adapted coordinates $(q^1,q^2,u)$ gives
\begin{equation}
\Delta_{\mathbb{R}^3} = \Delta_{LB} + \partial_u^2 - (H^2-K) + \mathcal{O}(u).
\end{equation}
After restricting to the transverse ground state and eliminating the $u$-direction,
    the effective operator contains the additional curvature term.

Thus the discrepancy arises because the Laplacian depends on the full embedding 
    metric, and eliminating the normal degree of freedom after quantization 
generates curvature contributions absent in the purely algebraic Dirac reduction.
\end{proof}

\subsection{Symplectic Reduction Interpretation}

The torus may also be understood from the viewpoint of symplectic reduction. 
Consider a free particle in $T^*\mathbb{R}^3$ with canonical symplectic form 
$\omega = d p_i \wedge d x^i$. Let $\Sigma \subset \mathbb{R}^3$ be the 
embedded torus defined implicitly by a constraint function $F(\mathbf{x})=0$.

The holonomic constraint $F(\mathbf{x})=0$ together with the consistency condition
$\nabla F \cdot \mathbf{p}=0$ defines a second-class constraint surface in 
phase space. The Dirac bracket projects momenta to the tangent bundle $T\Sigma$ 
and induces a reduced symplectic form
\begin{equation}
\omega_{\Sigma} = \iota^* \omega,
\end{equation}
where $\iota: T^*\Sigma \hookrightarrow T^*\mathbb{R}^3$ is the embedding.

Equivalently, one may describe the reduction as follows. Let $u$ denote the 
normal coordinate to the surface. Translations along the normal direction 
generate a Hamiltonian flow with moment map $\mu = p_u$. The constrained 
dynamics corresponds to fixing $\mu=0$ and quotienting by this normal flow.
The reduced phase space is therefore
\begin{equation}
\mu^{-1}(0)/\mathbb{R} \cong T^*\Sigma.
\end{equation}

Thus the torus system arises by symplectic reduction of $T^*\mathbb{R}^3$ with
respect to the normal translation symmetry. The curvature-induced potential does
not appear at the classical level; it is generated purely by the quantum 
reduction of the Laplacian under elimination of the transverse degree of freedom.

From this perspective, the geometric potential measures the discrepancy between
classical symplectic reduction and quantum reduction of the kinetic operator. 
The term $(H^2-K)$ is therefore a quantum anomaly associated with embedding 
curvature.

\subsection*{Marsden--Weinstein vs Dirac Reduction}

Marsden--Weinstein reduction applies to a symplectic manifold $(M,\omega)$ 
equipped with a Hamiltonian action of a Lie group $G$ with moment map 
$\mu: M \to \mathfrak{g}^*$. For a regular value $0$, the reduced phase 
space is the symplectic quotient
\begin{equation}
M//G = \mu^{-1}(0)/G.
\end{equation}
This procedure removes first-class constraints generated by symmetries 
and produces a new symplectic manifold whose dimension is reduced by $2\dim G$.

By contrast, Dirac reduction treats second-class constraints $\phi_i=0$ 
whose Poisson brackets form an invertible matrix. No symmetry group is 
quotiented out. Instead, the Poisson bracket is replaced by the Dirac bracket
\begin{equation}
{A,B}_D = {A,B} - {A,\phi_i}C^{ij}{\phi_j,B},
\end{equation}
thereby projecting dynamics onto the constraint surface directly. The reduced 
space inherits a symplectic structure but not as a quotient by a 
global compact symmetry.

In the surface-constrained particle problem, the elimination of the normal 
coordinate corresponds to second-class constraints rather than a compact 
Hamiltonian group action. Therefore the reduction mechanism is of Dirac type, 
not Marsden--Weinstein type.

\subsection*{Historical Remarks}

The principle that ``quantization commutes with reduction'' was formulated 
and proven in the context of geometric quantization by Guillemin and Sternberg 
in the early 1980s. Their theorem established that, under suitable compactness 
and regularity assumptions,
\begin{equation}
Q(M)^G \cong Q(M//G),
\end{equation}
meaning that taking $G$-invariant quantum states before reduction is equivalent 
to quantizing the reduced symplectic quotient.

The surface constraint problem  lies outside that framework because it involves 
second-class constraints and operator-level projection rather than reduction by a 
compact symmetry group. The curvature-induced geometric potential therefore 
reflects a distinct mechanism from the Marsden--Weinstein setting.

The above theorem appears at first sight to contradict the celebrated principle 
of Guillemin and Sternberg that ``quantization commutes with reduction.'' In their 
framework, one considers a compact symplectic manifold $(M,\omega)$ equipped with a 
Hamiltonian action of a compact Lie group $G$ and corresponding moment map $\mu$. 
Under suitable hypotheses (properness, compactness, existence of a prequantum 
line bundle), geometric quantization satisfies
\begin{equation}
Q(M)^G \cong Q(M//G),
\end{equation}
where $M//G = \mu^{-1}(0)/G$ is the symplectic quotient.

The apparent tension with the thin-layer result is resolved by observing that 
the constrained-surface problem does not fit the hypotheses of the 
Guillemin--Sternberg theorem:
\begin{itemize}
\item The ambient phase space $T^*\mathbb{R}^3$ is noncompact.
\item The constraint eliminating the normal direction is second-class rather 
    than generated by a compact symmetry group.
\item The reduction is implemented by projection onto an embedded submanifold 
    rather than quotient by a compact Hamiltonian group action.
\end{itemize}

In particular, the thin-layer procedure removes a transverse degree of freedom 
by spectral projection onto a bound state of a confining potential. This operation
is not a symplectic quotient in the strict Marsden--Weinstein sense, but rather 
a limit of operator-level projections. Consequently, the resulting quantum 
Hamiltonian retains curvature information encoded in the embedding metric.

Thus there is no contradiction: the Guillemin--Sternberg theorem applies to 
reduction by compact symmetry groups at the level of geometric quantization,
whereas the surface constraint considered here corresponds to elimination of 
second-class constraints and does not arise from such a group action. The 
curvature-induced term $(H^2-K)$ therefore reflects the different mathematical 
nature of the reduction procedure.

