\chapter{Quantization }
\label{qnt}

Quantization is the collection of procedures by which a classical mechanical 
system is associated with a non-relativistic quantum theory. 
Naturally  quantization is neither unique nor algorithmic in general 
\footnote{ Logically we should speak of de-quantization --- transition from quantum to classical system }, 
and no single scheme applies uniformly to all classical systems.Below  we analyze some quantization schemes, 
discuss no-go theorems limiting the existence of consistent quantization maps.

\section {Canonical Quantization}
\label{canonical-quantization}

Canonical quantization assigns self-adjoint operators $\hat q^i, \hat p_j$ on Hilbert space $ \mathcal{H} $
satisfying canonical commutation relations
\begin{equation}
[\hat q^i, \hat p_j] = i\hbar , \delta^i{}_j.
\end{equation}
to classical observables $q^i, p_j$ --- coordinates on phase space $ \mathbb{R}^{2 n} $.
Though this prescription is insufficient for general systems and obscures deeper structural issues: 
the relation between classical observables and quantum operators, the (non-)existence of functorial 
quantization procedures, and the origin of inequivalent quantizations. 

For most practical problems it is realized as
\begin{equation}
q^i \mapsto \hat q^i = q^i, \qquad p_j \mapsto \hat p_j = -i\hbar \frac{\partial}{\partial q^j},
\end{equation}
acting on $\mathcal{H}=L^2(Q,dq)$. For general Hamiltonians, operator ordering ambiguities arise, 
and additional analytic input is required to ensure self-adjointness and well-defined dynamics.
The standard name "canonical quantization" is misleading.

\section{Dirac Quantization}
\label{dirac-quantization}

A classical non-relativistic system is described by a symplectic manifold $(M,\omega)$, 
typically $M=T^*Q$ with canonical symplectic form
\begin{equation}
\omega = dq^i \wedge dp_i.
\end{equation}
Observables are smooth functions $f \in C^\infty(M)$ equipped with the Poisson bracket
\begin{equation}
{f,g} = \omega(X_f,X_g),
\end{equation}
where $X_f$ is the Hamiltonian vector field defined by $\iota_{X_f}\omega = df$. 
The resulting Poisson algebra encodes both kinematics and dynamics.

Time evolution is generated by a Hamiltonian function $H$ according to
\begin{equation}
\frac{df}{dt} = {f,H}.
\end{equation}
This structure motivates the canonical correspondence between Poisson brackets and 
commutators in quantum theory.

One often idealizes quantization as a map
\begin{equation}
Q : C^\infty(M) \to \mathcal{A}(\mathcal{H}),
\end{equation}
from classical observables to operators on a Hilbert space $\mathcal{H}$, satisfying 
linearity, reality, normalization, and the Poisson--commutator correspondence
\begin{equation}
Q({f,g}) = \frac{1}{i\hbar}[Q(f),Q(g)].
\end{equation}

However the Groenewold--Van Hove theorem shows that no such map exists on the full 
Poisson algebra of observables for generic phase spaces. 
Even for $M=\mathbb{R}^{2n}$, the correspondence can be consistently imposed only on 
restricted subalgebras \footnote{ In my paper \cite{panfil-q} I studied these subalgebras},
demonstrating that quantization is intrinsically limited.

\section{Geometric Quantization}


Geometric quantization begins with a complex line bundle $L \to M$ equipped with a connection $\nabla$ satisfying
\begin{equation}
\mathrm{curv}(\nabla) = \frac{1}{\hbar}\omega.
\end{equation}
This requires the integrality condition $[\omega]/2\pi\hbar \in H^2(M,\mathbb{Z})$.


To obtain a physical Hilbert space, one introduces a polarization selecting a subset of compatible observables. The choice of polarization is generally non-unique and may not exist globally.

\section{Deformation Quantization}

Deformation quantization replaces the pointwise product of functions on phase space by a star product
\begin{equation}
f \star g = fg + \frac{i\hbar}{2}{f,g} + O(\hbar^2).
\end{equation}
Quantum mechanics is then viewed as a deformation of the classical observable algebra rather than as an operator theory ab initio.

\section{Path Integral Quantization}

In the path-integral approach, transition amplitudes are formally expressed as
\begin{equation}
\langle q_f,t_f | q_i,t_i \rangle = \int \mathcal{D}q, e^{\frac{i}{\hbar}S[q]},
\end{equation}
where $S$ is the classical action. While powerful and intuitive, this formulation typically requires auxiliary constructions for mathematical rigor.

\section{Stochastic quantization}
\label{stochastic-quantization}

\section{Tomographic Quantization}
\label{tomographic-quantization}

