\section{ Deep Structural View}\label{deep-structural-view}
Dirac notation conflates three levels:
\textbar{} Level \textbar{} Meaning of \(\langle a|A|b\rangle\)
\textbar{}\\
\textbar{} Linear algebra \textbar{} Matrix element \textbar{}\\
\textbar{} Hilbert space \textbar{} Inner product \textbar{}\\
\textbar{} Rigged Hilbert space \textbar{} Distribution pairing
\textbar{}\\
\textbar{} Spectral theory \textbar{} Integral against PVM \textbar{}\\
\textbar{} C*-algebra \textbar{} State applied to operator \textbar{}
Same symbol --- different mathematical structures.
That is the source of ambiguity.
\section{ Geometric Interpretation}\label{geometric-interpretation}
In geometric formulation:
\begin{itemize}
\item
  States are rays in projective Hilbert space.
\item
  Observables are functions via moment map.
\item
  Expectation value:
\end{itemize}
\[\langle \psi|A|\psi\rangle\]
is a Hamiltonian function on \(\mathbb P(\mathcal H)\).
But:
\[\langle a|A|b\rangle\]
is not geometrically intrinsic --- it depends on representatives in the
Hilbert bundle.
Thus it is gauge-dependent (phase dependent).
\section{9️⃣ Final Answer}\label{final-answer}
Is it ambiguous?
\subsubsection{Mathematically:}\label{mathematically}
Yes --- unless domains and spaces are specified.
\subsubsection{Physically:}\label{physically}
Usually no --- physicists assume everything is well-behaved.
\subsubsection{Conceptually:}\label{conceptually}
It compresses multiple structures into one symbol.
Dirac notation trades precision for expressive power.
