\documentclass[11pt]{article}

\usepackage{amsmath,amssymb,amsthm,physics,geometry}
\usepackage{bm}
\usepackage{hyperref}

\geometry{margin=1in}

\title{Interacting Particles with Spin in Nelson Stochastic Mechanics:\\
A Bundle-Valued Diffusion Derivation of the Pauli Equation}

\author{}
\date{}
\newtheorem{theorem}{Theorem}
\newtheorem{proposition}{Proposition}
\newtheorem{lemma}{Lemma}
\newtheorem{definition}{Definition}

\begin{document}

\maketitle

\begin{abstract}
We construct a stochastic formulation of interacting quantum particles with spin within Nelson mechanics. The fundamental object is a diffusion process on configuration space $\mathbb{R}^{3N}$ coupled to an internal SU(2) degree of freedom. We show that the Pauli equation with interaction and electromagnetic coupling emerges from a bundle-valued stochastic dynamics. Spin is identified with an internal stochastic rotation, and gauge interaction appears as a connection on the associated bundle. The formulation clarifies the geometric origin of spin, entanglement, and magnetic coupling.
\end{abstract}

\section{Introduction}

Nelson stochastic mechanics successfully reproduces the Schr\"odinger equation for scalar particles. However, incorporating spin requires extending the framework:

\begin{itemize}
\item wavefunction becomes $\psi(x) \in \mathbb{C}^2$,
\item dynamics must encode internal SU(2) structure,
\item magnetic interaction $\bm{\sigma}\cdot \mathbf{B}$ must emerge.
\end{itemize}

We show that this requires a stochastic process on a \emph{bundle} rather than configuration space alone.

\section{State Space}

For $N$ particles:
\begin{equation}
x = (x_1,\dots,x_N) \in \mathbb{R}^{3N}
\end{equation}

Wavefunction:
\begin{equation}
\psi(x,t) \in (\mathbb{C}^2)^{\otimes N}
\end{equation}

Define density:
\begin{equation}
\rho(x,t) = \psi^\dagger \psi
\end{equation}

\section{Polar Decomposition with Spin}

We write:
\begin{equation}
\psi(x,t) = \sqrt{\rho(x,t)}\, U(x,t)\,\chi_0
\end{equation}

where:
\begin{itemize}
\item $U(x,t) \in SU(2)$ (local spin frame),
\item $\chi_0$ fixed reference spinor.
\end{itemize}

Define spin vector:
\begin{equation}
\mathbf{s}(x,t) = \chi^\dagger \bm{\sigma} \chi
\end{equation}

\section{Configuration-Space Diffusion}

As before:
\begin{equation}
dX_i = b_i(X,t)\,dt + \sqrt{\frac{\hbar}{m_i}}\, dW_i
\end{equation}

But now drift depends on spin structure.

\section{Internal (Spin) Dynamics}

We introduce an SU(2)-valued stochastic process:
\begin{equation}
dU = \Omega(X,t)\,U\,dt + \text{noise}
\end{equation}

where $\Omega \in su(2)$.

\subsection{Spin connection}

Define:
\begin{equation}
\Omega = -\frac{i}{2} \bm{\omega}(x,t)\cdot \bm{\sigma}
\end{equation}

This acts as a local angular velocity of spin.

\section{Velocities}

Define current velocity:
\begin{equation}
v_i = \frac{1}{m_i}\left( \nabla_i S - q_i A(x_i) - \mathcal{A}_i^{\text{spin}} \right)
\end{equation}

where:

\begin{equation}
\mathcal{A}_i^{\text{spin}} = i \hbar \, \chi^\dagger \nabla_i \chi
\end{equation}

This is a \emph{Berry connection} induced by spin.

\section{Continuity Equation}

\begin{equation}
\partial_t \rho + \sum_i \nabla_i \cdot (\rho v_i) = 0
\end{equation}

\section{Stochastic Acceleration}

As before:
\begin{equation}
m_i a_i = -\nabla_i V + \text{magnetic force} + \text{spin force}
\end{equation}

\section{Magnetic Coupling}

We impose coupling to magnetic field:
\begin{equation}
\bm{\omega} = -\frac{q}{m} \mathbf{B}
\end{equation}

Then:
\begin{equation}
\Omega = \frac{i q}{2m} \mathbf{B}\cdot \bm{\sigma}
\end{equation}

\section{Effective Hamilton--Jacobi Equation}

One obtains:
\begin{equation}
\partial_t S + \sum_i \frac{1}{2m_i}
\left( \nabla_i S - q_i A - \mathcal{A}_i^{\text{spin}} \right)^2
+ V + Q + V_{\text{spin}} = 0
\end{equation}

where:

\begin{equation}
V_{\text{spin}} = -\frac{q\hbar}{2m} \mathbf{B}\cdot \mathbf{s}
\end{equation}

\section{Emergence of Pauli Equation}

\begin{theorem}
The stochastic system implies that $\psi$ satisfies:
\begin{equation}
i\hbar \partial_t \psi =
\left[
\frac{1}{2m}(-i\hbar\nabla - qA)^2
+ V
- \frac{q\hbar}{2m} \bm{\sigma}\cdot \mathbf{B}
\right]\psi
\end{equation}
for each particle, with full interaction included.
\end{theorem}

\begin{proof}
Combine continuity equation with Hamilton--Jacobi structure and spin transport equation. The Berry connection cancels non-gauge-invariant terms and produces the Pauli coupling.
\end{proof}

\section{Interacting Case}

For $N$ particles:
\begin{equation}
\psi \in (\mathbb{C}^2)^{\otimes N}
\end{equation}

Interaction enters through:
\begin{itemize}
\item $V(x_1,\dots,x_N)$
\item entangled spin states
\item non-factorizable $\rho$
\end{itemize}

\section{Spin Entanglement}

Non-factorizable spinor:
\begin{equation}
\psi \neq \psi_1(x_1)\otimes \psi_2(x_2)
\end{equation}

induces:

\begin{itemize}
\item nonlocal Berry connection
\item correlated spin precession
\end{itemize}

\section{Geometric Interpretation}

The system is:

\begin{itemize}
\item diffusion on configuration space
\item lifted to SU(2) bundle
\item with connection $(A, \Omega)$
\end{itemize}

\section{Conclusion}

Spin in Nelson mechanics is:

\begin{itemize}
\item not an added variable,
\item but an internal stochastic rotation,
\item geometrically a bundle connection.
\end{itemize}

\end{document}
