\documentclass[russian,english]{book}            
\usepackage{babel}
\usepackage{csquotes}
\usepackage[backend=biber,style=numeric,sorting=none]{biblatex}
\usepackage{amsmath,amssymb,amsfonts}
%\usepackage{amsmath}
%\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{fontspec}
\usepackage{tikz}
%\usepackage{geometry}
\usepackage{physics}
\usepackage{bm}
\usepackage{tensor}
\usepackage{mathtools}
%\usetikzlibrary{quotes}
%\usepackage{verse}
\usepackage{epigraph}
\usepackage{siunitx}
\defaultfontfeatures{Ligatures={TeX}} 
% for russian letters
\setmainfont{DejaVu Serif} 
\setsansfont{DejaVu Sans} 
\setmonofont{DejaVu Sans Mono} 
\addbibresource{top.bib}
\newtheorem{theorem}{Theorem}
\newtheorem{definition}{Definition}
\newtheorem{proposition}{Proposition}
\newtheorem{remark}{Remark}
\newtheorem{axiom}{Axiom}

%\includeonly{quant-mech}

\title{Thoughts on Physics}
\author{Shimon Panfil}


\begin{document}  

\maketitle

\frontmatter
\chapter{Preface}

\epigraph{Physics is becoming so unbelievably complex that it is taking longer
and longer to train a physicist. It is taking so long, in fact, to train a 
physicist to the place where he understands the nature of physical problems that 
he is already too old to solve them.}{Eugene Paul Wigner}

 I am physicist. I've got my MSc in Physics from Novosibirsk State University in 1977 
 and PhD in Theoretical and Mathematical Physics from Institute of Nuclear Physics 
 in 1984 (both in Novosibirsk, Russia). Still in Novosibirsk I enjoyed doing 
 theoretical physics for 15 years. In 1992 I made aliyah (repatriated to Israel)
 and my academic career ended. I never regretted both events.

During next 30 years I was "industrial physicist",engineer,algorithm designer, 
software developer and even quant for a couple of years (from 2007 till 2009).
The latter was actually failure, being physicist I tried to understand how all this
is working and could not. Fortunately 2008 crisis came to the rescue: it does not
work at all.

For me being physicist means to always try to understand the problem, to 
construct mathematical model, investigate it and compare the results to 
real world. This approach contradicts very popular and sometimes successful
pure statistical ones, like AI and all that. It is hard to miss bitter irony
in the fact that the Nobel Prize in Physics in 2024 was awarded "for foundational discoveries 
and inventions that enable machine learning with artificial neural networks".

Now I am retired and decided to share my thoughts about  physics
in the book, which is a set of loosely connected comments on questions 
which frequently arise on studying theoretical physics but rarely answered.
 These questions may be not so important for calculations, but essential for understanding.
 In a sense it is feedback of  industrial physics to theoretical physics. As young
 theoretical physicist I always tried to study the problem in its most general form,
 never made numerical calculations being convinced that order of magnitude
 estimations are enough. Now I understand the importance of proper balance between
 generality and practicality, and always ask myself how small (large) 
 is small (large) enough.
  Partially this book is motivated by the fact that I see no significant progress in 
 quantum field theory and theory of elementary particles for the last 50 years!  

 Probably the world would be better without this book.
 Anyway since I don't intend to publish it on paper, the harm will
be minimal.

\tableofcontents                        

\mainmatter

%\input{intro.tex}
\include{intro}

%\input{class-mech}
\include{class-mech}

%\input{class-ed}
\include{class-ed}

%\input{relativity}
\include{relativity}

\include{quant-mech}

\backmatter
\printbibliography[
heading=bibintoc,
title={References}
] 
\clearpage
\end{document}
